---
title: "Configuration Matrices"
author: "Nils Reiter, nils.reiter@ims.uni-stuttgart.de"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    fig_width: 6 
    fig_height: 5
  md_document: 
    variant: markdown_github
    toc: true
    includes:
      before_body:
      - version.md
      - vig-Configuration-Matrices.md
vignette: >
  %\VignetteIndexEntry{Configuration matrices and co-occurrence networks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
---

```{r, include=FALSE}
require(DramaAnalysis)
v <- packageVersion("DramaAnalysis")
```


## Matrices

As configuration matrices [@Pfister:1988aa] are dependent on information about acts and scenes, we need to load the texts in such a way that this information is present. The function `load.text2()` does exactly that.  Alternatively, one can use the dataset `rksp.0.mtext`.


```{r setup}
require(DramaAnalysis)
data(rksp.0)

colnames(rksp.0$mtext)
```

In addition to the regular text and speakers, this table also contains informatio about the scenes and acts in which each token is spoken.

```{r}
c <- configuration(rksp.0$mtext)
c$matrix
c$figure
```

This creates a basic configuration matrix, but instead of just containing the presence or absence of a figure, it contains the number of spoken tokens for each act for each figure. 

We can use this easily to create a stacked bar chart, showing the distribution visually (you probably need a color palette with more than ten colors ... )

```{r}
par(mar=c(2,2,2,10))
barplot(c$matrix, 
        legend.text = c$figure, # legend text
        args.legend = list(cex=0.5, # legend font size
                           x=7.5, # legend x position
                           y=max(colSums(c$matrix)) # legend y pos
                        ), 
        col=qd.colors)
```

Since each act has a different length, it would be useful to normalize each block, according to the total number of spoken tokens. This way, we can display the relative active presence of each figure in each act. We normalize by dividing by the sum of each column.

```{r}
c$matrix <- scale(c$matrix, center=FALSE, scale=colSums(c$matrix))
c$matrix
par(mar=c(2,2,2,10))
barplot(c$matrix, 
        legend.text=c$figure, # set legend text
        args.legend = list(cex=0.5, # legend font size
                           x=7.5, # legend x position
                           y=max(colSums(c$matrix)) # legend y pos
                        ),
        col=qd.colors)
```

## Copresence

Configuration matrices are also often used to get an overview of who is copresent on stage. First, we create a configuration matrix that only represents presence or absence of a figure (and we switch to scenes). Obviously, the resulting matrix has many more columns.
```{r}
c <- configuration(rksp.0$mtext, onlyPresence = TRUE, by="Scene")

# to see the matrix (not shown here):
# c$matrix 
``` 

Creating a co-occurrence matrix is a simple matter of matrix multiplication:
```{r}
rksp.0.co <- c$matrix %*% t(c$matrix)
rksp.0.co

# add figure names
rownames(rksp.0.co) <- c$figure
colnames(rksp.0.co) <- c$figure
```

### As Heatmap

This can be visualised in a simple heat map:
```{r}
# since it's a square matrix, we don't need the bottom right triangle
# and diagonales.
rksp.0.co[lower.tri(rksp.0.co,diag=TRUE)] <- NA

par(mar=c(10,10,1,1)) # plot margins
image(rksp.0.co, 
      col = rgb(64,111,184, alpha=(seq(0,255)),
                maxColorValue = 256),
      xaxt= "n",  # no x axis
      yaxt= "n",  # no y axis
      frame=TRUE  # print a frame around the heatmap
      )
# add the x axis
axis(1, at = seq(0,1,length.out = length(c$figure)), labels = c$figure, las=3)
# add the y axis
axis(2, at = seq(0,1,length.out = length(c$figure)), labels = c$figure, las=1)
```

Apparently, Marinelli and Der Prinz have the most shared scenes. Marinelli also shares a scene with most other figures (vertical bar).

### As Network

... or even a co-occurrence network, using the package `igraph`. A nice introduction in igraph can be found in [@Arnold:2015aa], particularly for literary networks.


```{r message=FALSE}
require(igraph)
```

Technically, the matrix we created before is an [adjacency matrix](https://en.wikipedia.org/wiki/Adjacency_matrix). It is therefore simple to convert it to a graph, and igraph offers the function `graph_from_adjacency_matrix()` for this.

```{r}
g <- graph_from_adjacency_matrix(rksp.0.co, 
                                 weighted=TRUE,     # weighted graph
                                 mode="undirected", # no direction
                                 diag=FALSE         # no looping edges
                                )

# Now we plot
plot.igraph(g, 
            layout=layout_with_gem,       # how to lay out the graph
            main="Co-Occurrence Network: Emilia Galotti",  # title
            vertex.label.cex=0.6,         # label size
            vertex.label.color="black",   # font color
            vertex.color=qd.colors[4],    # vertex color
            vertex.frame.color=NA,        # no vertex border
            edge.width=E(g)$weight        # scale edges according to their weight
            )  
```

### Graph Export

As a final step, one might want to further work on the graph using [Gephi](https://gephi.org), or other tools. In order to do so, one can export the graph into an appropriate file:
```{r eval=FALSE}
write_graph(g, 
            "rksp.0.graphml",
            format="graphml")
```

This results in a file called [`rksp.0.graphml`](rksp.0.graphml), that starts similarly as this:
```xml
<?xml version="1.0" encoding="UTF-8"?>
<graphml xmlns="http://graphml.graphdrawing.org/xmlns"
         xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance"
         xsi:schemaLocation="http://graphml.graphdrawing.org/xmlns
         http://graphml.graphdrawing.org/xmlns/1.0/graphml.xsd">
<!-- Created by igraph -->
  <key id="v_name" for="node" attr.name="name" attr.type="string"/>
  <key id="e_weight" for="edge" attr.name="weight" attr.type="double"/>
  <graph id="G" edgedefault="undirected">
    <node id="n0">
      <data key="v_name">DER KAMMERDIENER</data>
    </node>
    <node id="n1">
      <data key="v_name">DER PRINZ</data>
    </node>
    ...
```

This file can be opened with Gephi.

## References