---
title: "Figure Statistics"
author: "Nils Reiter"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    fig_width: 6
    fig_height: 5
  md_document:
    variant: markdown_github
    toc: true
    includes:
      before_body:
      - version.md
      - vig-Figure-Statistics.md


vignette: >
  %\VignetteIndexEntry{Who speaks when how much and often?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(DramaAnalysis)
require(data.table)

v <- packageVersion("DramaAnalysis")
```


We're assuming here that we have loaded some texts using `loadText()`, and that this text is stored as a data.frame in the variable `text`. For this example, we have loaded Emilia Galotti and Romeo und Julia, both coming from the TextGrid repository. For demo purposes, one can use the data sets included in the package.

```{r}
# Load Emilia Galotti
data(rksp.0)

# Load Miß Sara Sampson
data(rjmw.0)

text <- rbind(rksp.0$mtext, rjmw.0$mtext)

```

## Who's talking how much? (Tokens)


First, we calculate figure statistics. 
```{r}
fstat <- figureStatistics(text, names=TRUE, normalize=FALSE)
summary(fstat)
```
This figure statistics table already contains all the information we need and can be inspected and analysed directly. The remaining steps are only needed if you want to plot this information as a stacked bar chart.

```{r}
# Create a matrix
mat <- figurematrix(fstat)
summary(mat)
head(mat$values,5)
```

`mat` is a list containing three things: `values` is a matrix storing the number of tokens a figure speaks, `labels` contains the names of figures corresponding to the respective cell in `values` and `cs` contains summed token numbers, such that the entries can be stacked.

```{r}
# Plot a stacked bar plot
b <- barplot(mat$values,col=qd.colors)

# Add figure names (if needed/wanted)
text(x=b,y=t(mat$cs+(mat$values/2)),
     labels=t(substr(mat$labels,0,20)))
```

Since many of these labels are barely readable, we can restrict the number of displayed labels to, say, the top 5 figures (i.e., the five figures that speak the most). We do this by selecting only the top 5 rows in the matrices, using the function `head()`.

```{r}
# Plot a stacked bar plot
b <- barplot(mat$values,col=qd.colors)

top <- 5

# Add figure names (if needed/wanted)
text(x=b, y=t(head(mat$cs,top)+(head(mat$values,top)/2)),
     labels=t(substr(head(mat$labels,top),0,20)))
```

## Who's talking how often? (Utterances)

So far, we have counted words. Now we will turn to utterances, and their properties.

First, we will use the function `utteranceStatistics()` to extract numbers about utterances
```{r}
ustat <- utteranceStatistics(rksp.0$mtext, 
                              numberOfFigures = 10, # include 10 figures per drama
                              normalizeByDramaLength = FALSE # use absolute values
                            )
colnames(ustat)
```

This creates a table with a single row for each utterance. We can now inspect the variance in utterance length

```{r}
par(mar=c(10,2,2,2))
boxplot(utteranceLength ~ figure, # what do we want to correlate 
        data=ustat, 
        las = 3 # rotate axis labels
      )
```

## When are figures talking?

While the above displays the *length* of utterances, we can also display the position of utterances, i.e., where in the text are they taking place?

```{r}
par(mar=c(2,10,2,2))
stripchart(begin ~ figure, data=ustat, 
           las=1, # horizontal labels
           pch=20, # use a small bullet as symbol
           col=qd.colors # get nice colors
           )
```

Each dot in this plot represents one utterance, the x-axis is measured in character positions. This is not really intuitive, but the flow from left to right represents the flow of the text.

Now it would be useful to include information on act/scene boundaries in this plot. This can be done by accessing the segmented text table. The first  commands work just as before, with the exception of being applied to the variable `rksp.0.mtext`.

```{r}

ustat <- utteranceStatistics(rksp.0$mtext, 
                              numberOfFigures = 10, # include 10 figures per drama
                              normalizeByDramaLength = FALSE # use absolute values
                            )
par(mar=c(2,10,2,2))
stripchart(begin ~ figure, data=ustat, 
           las=1, # horizontal labels
           pch=20, # use a small bullet as symbol
           col=qd.colors, # get nice colors
           xaxt="n" # suppress the x axis
           )

# add vertical lines for act beginning
abline(v=unique(rksp.0$mtext$begin.Act)[-1])
```

Please note that the information contained in this plot is very similar to the information in the [visual and relative configuration matrices](Configuration-Matrices).

## When are characters mentioned?

When characters are speaking on stage, they are actively present. But they can also be passively present, if other characters refer to them. Both levels of presence can be extracted with the `presence()` function:
```{r}
pres <- presence(rksp.0$mtext)
pres
```

As we can see, each character has a few numbers associated: The column `actives` shows the number of scenes in which the character is actively present. This is equivalent to the information in the configuration matrix. The column `passives` shows the number of scenes in which a character is mentioned. By default, this excludes the scenes in which they are present themselves (this behaviour can be changed by adding the parameter `passiveOnlyWhenNotActive = TRUE` to the call of the `presence` function). 

A simple visualisation that shows the characters active and passive presence in one plot can be generated like this: The first line (`plot()`) is responsible for the plotting of the symbols, the second line (`text()`) adds the character names or ids numbers.

```{r}
plot(x=pres$active/pres$scenes, 
     y=pres$passive/pres$scenes, 
     xlim=c(0,1), 
     ylim=c(0,1), 
     xlab="Active", 
     ylab="Passive",
     main="Character Presence")
text(x=pres$active/pres$scenes, 
     y=pres$passive/pres$scenes, 
     labels=substr(pres$figure,0,20), 
     pos=3)
```

