---
title: "Loading texts"
author: "Nils Reiter, nils.reiter@ims.uni-stuttgart.de"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    fig_width: 6
    fig_height: 5
  md_document:
    variant: markdown_github
    toc: true

vignette: >
  %\VignetteIndexEntry{Loading Texts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(DramaAnalysis)
```



In order to use the analysis functions in this package, you need NLP-pre-processed dramatic texts. To facilitate this, we are providing downloadable packages of pre-processed files in the correct format etc. Currently, we provide only the TextGrid repository as a collection, but others are planned. 

## Package Setup

First, you will need to find a directory in which the data files are to be stored. By default, a directory called `QuaDramA` will be created in your home directory (`~`). On Windows, this is known as the Documents-folder.

If you want to specify a different folder, you can do with the `setup()`-function:
```{r, eval=FALSE}
setup(dataDirectory = "YOUR_FOLDER")
```

The setup function is responsible for starting the Java Virtual Machine in the background, which we need in order to load the pre-processed texts. It also sets the option `qd.datadir` to the given directory, such that other functions know this. **Whenever the R session is restarted, this function needs to be called again.**

## Installation of Data

To install a data package, you simply run the function `installData()`. The download size is approximately 700MB, so it might take a while. Luckily, this only needs to be done once.

```{r, eval=FALSE}
installData()
```

In addition, we provide a categorization of these texts into genres. It can be downloaded in a similar way.

```{r, eval=FALSE}
installCollectionData()
```

## Loading of Dramatic texts

Each dramatic text in one of our data packages has an id, that we need to load the text. The id itself is just a sequence of alphanumeric symbols. The function `loadAllInstalledIds()` gives you a list of all installed ids. This is not particularly helpful, that's why we made an overview here.

