---
title: "Word Field Analysis"
author: "Nils Reiter, nils.reiter@ims.uni-stuttgart.de"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    fig_width: 6
    fig_height: 5
  md_document:
    variant: markdown_github
    toc: true
    includes:
      before_body:
      - version.md
      - vig-Word-Field-Analysis.md


vignette: >
  %\VignetteIndexEntry{Measuring speech topics using word fields}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
---

```{r, include=FALSE}
require(DramaAnalysis)
v <- packageVersion("DramaAnalysis")
```



A word field is basically of a list of words that belong to a common theme / topic / semantic group.

```{r include=FALSE}
require(DramaAnalysis)
```

## Word Field Definition
For **demo purposes** (this is really a toy example), we will define the word field of `Love` as containing the words `Liebe` and `Herz`.
In R, we can put them in a character vector:

```{r}
wf_love <- c("liebe", "herz")
```

We will test this word field on Romeo und Julia, which should be about love. 

```{r}
data(vndf.0.text)
```

## Single word field
The core of the word field analysis collecting statistics about a dictionary. Therefore, we use the function called `dictionaryStatisticsSingle()` (single, because we only want to analyse a single word field):

```{r}
dstat <- dictionaryStatisticsSingle(
  vndf.0.text,         # the text we want to process
  wordfield=wf_love,   # the word field
  names=TRUE,          # we want figure names instead of ids
  column="Token.lemma" # we count lemmas instead of surface forms
)

summary(dstat)
```

We can visualise these counts in a simple bar chart:

```{r fig.height=3}
# remove figures not using these words at all
dstat <- dstat[dstat$x>0,] 

barplot(dstat$x,                   # what to plot
        names.arg = dstat$figure,  # x axis labels
        las=3,                     # turn axis labels
        cex.names=0.6,             # smaller font on x axis
        col=qd.colors              # colors
      )
```

Apparently, Romeo and Juliet are mentioning these words a lot. 

Obviously we would want to **normalize** according to the total number of spoken words by a figure:
```{r}
dstat <- dictionaryStatisticsSingle(
  vndf.0.text,       # the text we want to process
  wordfield=wf_love, # the word field
  names=TRUE,        # we want figure names instead of ids
  normalizeByFigure = TRUE,   # apply normalization
  column = "Token.lemma"
)

# remove figures not using these words at all
dstat <- dstat[dstat$x>0,]

barplot(dstat$x, 
        names.arg = dstat$figure,  # x axis labels
        las=3,             # turn axis labels
        cex.names=0.8,     # smaller font on x axis
        col=qd.colors
      )
```


## Multiple Word Fields

The function `dictionaryStatistics()` can be used to analyse multiple dictionaries at once. To this end, it is assumed that dictionaries are stored in text files that either are stored locally or reachable via a URL. In any case, the parameter `baseurl` can be used to specify a prefix, to which the field names are appended. By default, the baseurl `https://github.com/quadrama/metadata/tree/master/fields/` is used. Appending the field name "Liebe", for instance, thus results in the URL `https://github.com/quadrama/metadata/tree/master/fields/Liebe`, which is a plain text file.

The following examples use the baseurl `https://raw.githubusercontent.com/quadrama/metadata/ec8ae3ddd32fa71be4ea0e7f6c634002a002823d/fields/`, i.e., a specific version of the fields we have been using in QuaDramA. 

```{r}

# code is more readable if we define this variable 
baseUrl <- "https://raw.githubusercontent.com/quadrama/metadata/ec8ae3ddd32fa71be4ea0e7f6c634002a002823d/fields/"


dstat <- dictionaryStatistics(
  vndf.0.text,  # the text
  fieldnames=  # fields we want to measure
    c("Liebe", "Krieg", "Familie", "Ratio", "Religion"),
  baseurl = baseUrl,            # specify url
  names = TRUE,                 # use figure names (instead of ids)
  normalizeByFigure = TRUE,   # normalization by figure
  normalizeByField = TRUE,    # normalization by field
  column = "Token.lemma"        # lemma-based stats
)
colnames(dstat)       
```

The variable `dstat` now contains multiple columns, one for each word field.

### Bar plot by figure

We can now display the distribution of the words of a single figure over these word fields:

```{r}
barplot(as.matrix(dstat[28,3:7]), # we select Romeo's line
        main="Romeo's speech",    # plot title
        col=qd.colors             # more colors
        )
```


### Bar plot by field

... but we can also analyse who uses  words of a certain field how often

```{r}
barplot(as.matrix(dstat$Liebe),
        main="Use of love words", # title for plot
        beside = TRUE,            # not stacked
        names.arg = dstat$figure, # x axis labels
        las=3,                    # rotation for labels
        col=qd.colors,            # more colors
        cex.names = 0.7           # font size
        )
```