---
title: "Word Field Analysis"
author: "Nils Reiter, nils.reiter@ims.uni-stuttgart.de"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    fig_width: 6
    fig_height: 5
  md_document:
    variant: markdown_github
    toc: true
    includes:
      before_body:
      - version.md
      - vig-Word-Field-Analysis.md


vignette: >
  %\VignetteIndexEntry{Measuring speech topics using word fields}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
---

```{r, include=FALSE}
require(DramaAnalysis)
v <- packageVersion("DramaAnalysis")
```



A word field is basically of a list of words that belong to a common theme / topic / semantic group.

```{r include=FALSE}
require(DramaAnalysis)
```

## Word Field Definition
For **demo purposes** (this is really a toy example), we will define the word field of `Love` as containing the words `Liebe` and `Herz`.
In R, we can put them in a character vector:

```{r}
wf_love <- c("liebe", "herz")
```

We will test this word field on Emilia Galotti, which should be about love. 

```{r}
data("rksp.0")
```

## Single word field
The core of the word field analysis collecting statistics about a dictionary. Therefore, we use the function called `dictionaryStatisticsSingle()` (single, because we only want to analyse a single word field):

```{r}
dstat <- dictionaryStatisticsSingle(
  rksp.0$mtext,         # the text we want to process
  wordfield=wf_love,   # the word field
  column="Token.lemma" # we count lemmas instead of surface forms
)

#summary(dstat)
```

We can visualise these counts in a simple bar chart:

```{r fig.height=3}
# remove figures not using these words at all
dstat <- dstat[dstat$x>0,] 

barplot(dstat$x,                   # what to plot
        names.arg = dstat$figure,  # x axis labels
        las=3,                     # turn axis labels
        cex.names=0.6,             # smaller font on x axis
        col=qd.colors              # colors
      )
```

Apparently, the prince is mentioning these words a lot. 

Obviously we would want to **normalize** according to the total number of spoken words by a figure:
```{r}
dstat <- dictionaryStatisticsSingle(
  rksp.0$mtext,       # the text we want to process
  wordfield=wf_love, # the word field
  normalizeByFigure = TRUE,   # apply normalization
  column = "Token.lemma"
)

# remove figures not using these words at all
dstat <- dstat[dstat$x>0,]

barplot(dstat$x, 
        names.arg = dstat$figure,  # x axis labels
        las=3,             # turn axis labels
        cex.names=0.8,     # smaller font on x axis
        col=qd.colors
      )
```


## Multiple Word Fields

The function `dictionaryStatistics()` can be used to analyse multiple dictionaries at once. To this end, dictionaries are represented as lists of character vectors. The (named) outer list contains the keywords, the vectors are just words associated with the keyword.

New dictionaries can be easily like this:
```{r}
wf <- list(Liebe=c("Liebe","Herz","Schmerz"), Hass=c("Hass","hassen"))
```

This dictionary contains the two entries `Liebe` (love) and `Hass` (hate), with 3 respective 2 entries. Dictionaries can be created in code, like shown above. In addition, the function `loadFields()` can be used to download dictionary content from a URL or a directory. By default, the function loads [this dictionary](https://github.com/quadrama/metadata/tree/master/fields/) from GitHub (that we used in publications), for the keywords `Liebe` and `Familie` (family). Since version 2.3.0, this dictionary is included in the package as `base_dictionary` and can be used right away. It is also the default dictionary used by `dictionaryStatistics()`.

```{r, eval=TRUE}
data("base_dictionary")
names(base_dictionary)
```

The function `loadFields()` offers parameters to load from different URLs via http or to load from plain text files that are stored locally. The latter can be achieved by specifying the directory as `baseurl`. Entries for each keyword should then be stored in a file named like the keyword, and ending with `txt` (by default, can be overwritten). See `?loadFields` for details. Some of the options can also be specified through `dictionaryStatistics()`, as exemplified below.

The following examples use the `base_dictionary`, i.e., a specific version of the fields we have been using in QuaDramA. 


```{r, eval=TRUE}

dstat <- dictionaryStatistics(
  rksp.0$mtext,  # the text
  fieldnames = # fields we want to measure (from base_dictionary)
    c("Liebe", "Familie", "Ratio", "Krieg", "Religion"),
  names = TRUE,                 # use figure names (instead of ids)
  normalizeByFigure = TRUE,   # normalization by figure
  normalizeByField = TRUE,    # normalization by field
  column = "Token.lemma"        # lemma-based stats
)
colnames(dstat)
```

The variable `dstat` now contains multiple columns, one for each word field.

### Bar plot by figure

We can now display the distribution of the words of a single figure over these word fields:

```{r}
barplot(as.matrix(dstat[9,4:8]), # we select Emilia's line
        main="Emilia's speech",    # plot title
        col=qd.colors             # more colors
        )
```


### Bar plot by field

... but we can also analyse who uses  words of a certain field how often

```{r}
barplot(as.matrix(dstat$Liebe),
        main="Use of love words", # title for plot
        beside = TRUE,            # not stacked
        names.arg = dstat[[3]],  # x axis labels
        las=3,                    # rotation for labels
        col=qd.colors,            # more colors
        cex.names = 0.7           # font size
        )
```

## Dictionary Based Distance

Technically, the output of `dictionaryStatistics()` is a  `data.frame`. This is suitable for most uses. In some cases, however, it's more suited to work with a matrix that only contains the raw numbers (i.e., number of family words). Calculating character distance based on dictionaries, for instance. For these cases, the function  offers a parameter `asList`. If it is set to `TRUE`, the function returns a list with four to six components. The first three components (`corpus`, `drama`, `figure`, `Number.Act`, `Number.Scene`) contain meta data, while the fourth component (`mat`) contains the values as a matrix.

You typically want to assign rownames to this matrix. If you're studying individual characters, their name might be a good idea.

```{r}
dsl <- dictionaryStatistics(rksp.0$mtext, 
                            fieldnames=c("Liebe", "Familie", "Ratio", "Krieg", "Religion"),
                            normalizeByFigure=TRUE,
                            asList=TRUE)
dsl$mat
```


Every character is now represent with five numbers, which can be interpreted as a vector in five-dimensional space. This means, we can easily apply distance metrics supplied by the function `dist()` (from the default package `stats`). By default, `dist()` calculates [Euclidean distance](https://en.wikipedia.org/wiki/Euclidean_distance). The data structure here is now similar to the one in the weighted configuration matrix, which means everything from [this vignette](Configuration-Matrices) can be applied here.

```{r}
cdist <- dist(dsl$mat)
cdist
```

In particular, we can show these distances in a network:

```{r}
require(igraph)
g <- graph_from_adjacency_matrix(as.matrix(cdist), 
                                 weighted=TRUE,     # weighted graph
                                 mode="undirected", # no direction
                                 diag=FALSE         # no looping edges
                                )

# Now we plot
plot.igraph(g, 
            layout=layout_with_fr,       # how to lay out the graph
            main="Dictionary distance network",  # title
            vertex.label.cex=0.6,         # label size
            vertex.label.color="black",   # font color
            vertex.color=qd.colors[5],    # vertex color
            vertex.frame.color=NA,        # no vertex border
            edge.width=E(g)$weight*100    # scale edges according to their weight 
                                          # (since the distances are so small, we multiply)
            )  
```

## Development over the course of the play

Since version 2.0 of this package, we can also count word fields by segment, i.e., by act or scene in a play. To do that, we add the parameter `segment="Act"` when we call the function `dictionaryStatistics()`. With the function `individualize()`, we can then re-organize this output so that each character has its own list.

```{r}
dsl <- dictionaryStatistics(rksp.0$mtext, 
                            fieldnames=c("Liebe", "Familie", "Ratio", "Krieg", "Religion"),
                            normalizeByFigure=TRUE,
                            asList=TRUE,
                            segment="Scene")

dsl <- regroup(dsl)

```

We can now easily plot the theme progress over the course of the play:

```{r}
matplot(dsl$marinelli$mat,type="l",col="black")
legend(x="topleft",legend=colnames(dsl$marinelli$mat),lty=1:5)
```

Or cumulatively add up the numbers:

```{r}
mat <- apply(dsl$marinelli$mat,2,cumsum)
matplot(mat,type="l",col="black")

# Add act lines
al <- rle(dsl$marinelli$Number.Act)
abline(v=cumsum(al$lengths[-1]))

legend(x="topleft",legend=colnames(mat),lty=1:5)

```
