---
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
  pdf_document:
    toc: true
    toc_depth: 2
params:
  id:
  - "turm:1"
  - "turm:2"
  - "turm:3"
  col: NULL
  dynamic: TRUE
  charAlignment: NULL
  topn: 10
---


```{r,echo=FALSE}
pver <- packageVersion("DramaAnalysis")
```


```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

if (is.null(.Options$qd.datadir)) { 
  DramaAnalysis::setup()
}

getYPositions <- function(x,b,limit=100) {
  rl <- apply(x, 2, function(xx) {
    xx <- as.double(xx)
    cs <- c(0,head(cumsum(xx),-1))
    r <- cs + xx/2
    r[xx<=limit] <- NA
    r  
  })
  rl
}

toFigureName <- function(x) {
  tools::toTitleCase(tolower(x))
}
library(data.table)
library(DramaAnalysis)
library(magrittr)
library(highcharter)
library(stringdist)

#print(class(params$id))
#print(params$id[[2]])
text <- loadSegmentedText(params$id)
meta <- loadMeta(params$id)
characters <- loadCharacters(params$id)
#titles <- lapply(meta,function(x) { as.character(x[1,"documentTitle"]) })

```


```{r}
if (is.null(params$charAlignment)) {
  
}
```

---
title: "QuaDramA Edition Comparison"

---

This report shows a number of quantitative properties extracted from several plays. 
The report assumes the plays to be different variants *of the same text* (what ever
that means).

# Meta data

```{r, results="asis"}


knitr::kable(meta,
             col.names = c("Corpus", "Text-Id", "Title", "Language", "Author name", "Author Id", "Translator name", "Translator Id", "Year written", "Year printed", "Year premiered", "Year translated"),
             format = "markdown")

```


# Structure {.tabset}



## Overview

```{r}
tbl <- text[,.(Acts=length(unique(begin.Act)),
               Scenes=length(unique(begin.Scene))),
            .(corpus,drama)]
rownames(tbl) <- paste(tbl$corpus,tbl$drama,sep=":")
tblm <- as.matrix(tbl[,3:ncol(tbl)])

barplot(tblm,
        beside=TRUE,
        legend.text = rownames(tbl),
        col=params$col)

#knitr::kable(tbl)

```

## Differences

# Characters {.tabset}

## Number of Characters {.tabset}

```{r}
tbl <- text[,.(Characters=length(unique(Speaker.figure_surface))),
            .(corpus,drama)]
rownames(tbl) <- paste(tbl$corpus,tbl$drama,sep=":")
tblm <- as.matrix(tbl[,3:ncol(tbl)])
```

### Overview

```{r}
tbl <- text[,.(Characters=length(unique(Speaker.figure_surface))),
            .(corpus,drama)]
rownames(tbl) <- paste(tbl$corpus,tbl$drama,sep=":")
tblm <- as.matrix(tbl[,3:ncol(tbl)])
rownames(tblm) <- rownames(tbl)
barplot(tblm,
        beside=TRUE,
        legend.text = rownames(tbl),
        col=params$col)

#knitr::kable(tbl)

```

### Table

#### Absolute numbers

```{r, results="asis"}
knitr::kable(tblm, format="markdown", row.names = TRUE)
```

#### Differences
```{r}
knitr::kable(as.matrix(dist(tblm)))
```

## Spoken words {.tabset}

### Overview

```{r}
tl <- limitFigures(text, threshold = params$topn)

fstat <- figureStatistics(tl, names=TRUE)

xt <- xtabs(tokens ~  drama+figure,data=fstat)

attr(xt, "class") <- NULL 
attr(xt, "call") <- NULL 
#barplot(xt, beside=TRUE,las=3)

xt <- as.data.frame(t(xt))

hc <- highchart()

hc <- hc_yAxis(hc, max=max(xt))
hc <- hc_xAxis(hc, categories=rownames(xt))

for (i in 1:ncol(xt)) {
  hc <- hc_add_series(hc, xt[,i], type="column", name=i)
}

hc

```

### Table

```{r, results="asis"}
fstat <- figureStatistics(text, names=TRUE)

xt <- xtabs(tokens ~  drama+figure,data=fstat)


knitr::kable(t(xt), format="markdown")


```


## Utterances {.tabset}

### Lengths

This plot shows the distribution of utterange lengths for the top *n* characters in each play. If two characters are aligned, their box is shown side by side. More on [box plots in Wikipedia](https://en.wikipedia.org/wiki/Box_plot). Dots outside of the box indicate potential outliers, color shows the play they belong to.

```{r}
#tl <- limitFigures(text, threshold = 15)
ustat <- utteranceStatistics(text, numberOfFigures = params$topn)

if (is.null(params$col)) {
  col = "black"
} else {
  col = params$col[1:length(params$id)]
}

hcboxplot(x=ustat$utteranceLength, 
          var=as.character(ustat$figure), 
          var2=as.character(ustat$drama),
          color=col
          )


```

### Table

```{r, results="asis"}

#b <- boxplot(utteranceLength~drama+figure, data=ustat, plot=FALSE)

#knitr::kable(ustat,format="markdown")
```


```{r, eval=FALSE}



utteranceDistances <- function(t) {
  tagg <- t[,.(text=paste0(Token.surface, collapse=" ")),.(corpus,drama,begin, Speaker.figure_surface)]
  pairs <- t(combn(nrow(tagg),2))
  for(i in 1:nrow(pairs)) {
    row_a <- pairs[i,1]
    row_b <- pairs[i,2]
    
  }
  
}

text[,
     mcp:=max(end),
     .(corpus,drama)][,
                      .(utteranceLength=.N,rbegin=begin/mcp),
                      .(corpus,drama,begin,Speaker.figure_surface)][]

```

# Presence and Utterances {.tabset}

```{r,echo=FALSE}
ustat <- utteranceStatistics(text, normalizeByDramaLength = FALSE, numberOfFigures = FALSE)
ustat$figure <- tools::toTitleCase(tolower(ustat$figure))
```

## All Utterances
```{r,echo=FALSE}
par(mar=c(1,9,1,0),xpd=FALSE)

for (id in params$ids) {
  plotUtterancePositions(ustat[paste(ustat$corpus,ustat$drama,sep=":") == id,],
                         text[paste(text$corpus,text$drama,sep=":") == id,],
                         xlab="")
}
```


# Configuration and Copresence {.tabset}

```{r}

```