---
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
  pdf_document:
    toc: true
    toc_depth: 2
params:
  id: "test:rksp.0"
  col: NULL
  topN: 10
  minTokens: 500
---

```{r,echo=FALSE}
pver <- packageVersion("DramaAnalysis")
```


```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

if (is.null(.Options$qd.datadir)) { 
  setup()
}

getYPositions <- function(x,b,limit=100) {
  rl <- apply(x, 2, function(xx) {
    xx <- as.double(xx)
    cs <- c(0,head(cumsum(xx),-1))
    r <- cs + xx/2
    r[xx<=limit] <- NA
    r  
  })
  rl
}

toFigureName <- function(x) {
  tools::toTitleCase(tolower(x))
}

library(magrittr)
library(highcharter)

mtext <- loadSegmentedText(params$id)
meta <- loadMeta(params$id)
characters <- loadCharacters(params$id)
title <- as.character(meta[1,"documentTitle"])

```

---
title: "QuaDramA Report: `r title`"

---

# Meta data

- Author(s): `r paste(paste0("[",meta$Name,"](http://d-nb.info/gnd/",meta$Pnd,")"),sep=", ")`
- Acts: `r length(unique(mtext$begin.Act))`
- Scenes: `r mtext[,length(unique(begin.Scene)),.(begin.Act)]$V1`

# Character Lists {.tabset}

## By Order of appearance

All figures, in the order they have their first appearance

```{r, echo=FALSE}
begins <- aggregate(mtext$begin, by=list(mtext$Speaker.figure_surface), min)
as.character(begins[order(begins$x),]$Group.1)
```

# Overall Speech Distribution {.tabset}


## Bar chart 

Relative amount of words spoken by each figure. 

```{r, echo=FALSE, messages=FALSE}
fstat <- figureStatistics(mtext, names=TRUE, normalize=FALSE)
mat <- figurematrix(fstat)
b <- barplot(mat$values,col=params$col)
mat$labels <- tools::toTitleCase(tolower(mat$labels))
top <- 15
text(x=b, y=t(head(mat$cs,top)+(head(mat$values,top)/2)),
     labels=t(substr(head(mat$labels,top),0,20)),cex=0.5)

```

## Table

```{r,echo=FALSE, results="asis"}
fstat <- figureStatistics(mtext, names=FALSE, normalize=FALSE)

table <- merge(fstat, characters, 
               by.x=c("corpus","drama","figure"),
               by.y=c("corpus","drama","figure_id"))
knitr::kable(table,format="pandoc")
```

# Presence and Utterances {.tabset}

## Top `r params$topN` Characters {.tabset}

This segment shows the top `r params$topN` characters. This number can be adjusted by supplying the parameter `topN` with a different value to the `report()` function.

```{r}
mtextl <- limitFigures(mtext, other=TRUE, threshold = params$topN);
ustatl <- utteranceStatistics(mtextl, normalizeByDramaLength = FALSE, numberOfFigures = FALSE)
```

### All Utterances
```{r}
ustatl$figure <- toFigureName(ustatl$figure)
par(mar=c(1,9,1,0),xpd=FALSE)
plotUtterancePositions(ustatl,mtextl,xlab="")
```

### Utterance Length Variation

```{r,echo=FALSE}
par(mar=c(2,9,0,0))
boxplot(utteranceLength ~ figure, data=ustatl, horizontal=TRUE,las=1,frame=FALSE)
```

## All Characters {.tabset}
```{r}
ustat <- utteranceStatistics(mtext, normalizeByDramaLength = FALSE, numberOfFigures = FALSE)
ustat$figure <- toFigureName(ustat$figure)
```

### All Utterances
```{r,echo=FALSE}
par(mar=c(1,9,1,0),xpd=FALSE)
plotUtterancePositions(ustat,mtext,xlab="")
```

### Utterance Length Variation

```{r,echo=FALSE}
par(mar=c(2,9,0,0))
boxplot(utteranceLength ~ figure, data=ustat, horizontal=TRUE,las=1,frame=FALSE)
```

# Configuration and Copresence {.tabset}

```{r}
actLimit <- 0.1
sceneLimit <- 0.3
```

This visualisation displays the relative amount of tokens uttered by a character. The plot includes names of characters only when they speak more than `r actLimit*100`% of the tokens in an act and more than  `r sceneLimit*100`% of the tokens in a scene.

## By Act 
```{r, echo=FALSE}
c <- configuration(mtext)

c$matrix <- scale(c$matrix, center=FALSE, scale=colSums(c$matrix))
rownames(c$matrix) <- c$figure

b <- barplot(c$matrix,
        col=params$col)



yPositions <- as.vector(getYPositions(c$matrix,b,limit=actLimit))
xPositions <- rep(b,each=nrow(c$matrix))
lab <- rep(as.character(c$figure),ncol(c$matrix))

xPositions <- xPositions[which(!is.na(yPositions))]
lab <- lab[which(!is.na(yPositions))]
yPositions <- yPositions[which(!is.na(yPositions))]

if (length(yPositions > 0)) {
  text(x=xPositions,y=yPositions,labels=tools::toTitleCase(tolower(lab)),cex=0.6)
} 
```

## By Scene

```{r, echo=FALSE}
c <- configuration(mtext, by="Scene")
c$matrix <- scale(c$matrix, center=FALSE, scale=colSums(c$matrix))
rownames(c$matrix) <- c$figure

b <- barplot(c$matrix,
        col=params$col)



yPositions <- as.vector(getYPositions(c$matrix,b,limit=sceneLimit))
xPositions <- rep(b,each=nrow(c$matrix))
lab <- rep(as.character(c$figure),ncol(c$matrix))

xPositions <- xPositions[which(!is.na(yPositions))]
lab <- lab[which(!is.na(yPositions))]
yPositions <- yPositions[which(!is.na(yPositions))]
if (length(yPositions > 0)) {
  text(x=xPositions,y=yPositions,labels=tools::toTitleCase(tolower(lab)),cex=0.4,srt=90)
}
```


#  Network Analysis  {.tabset}

## Copresence Network

In a copresence network, two vertices are connected if the characters they represent appear together on stage (= are co-present on stage). The network shown here is based on scenes, i.e., the thicker an edge, the more scenes the two connected characters are copresent.


```{r, echo=FALSE, message=FALSE}
c <- configuration(mtext, onlyPresence = TRUE, by="Scene")
co <- c$matrix %*% t(c$matrix)

# add figure names
rownames(co) <- toFigureName(c$figure)
colnames(co) <- toFigureName(c$figure)
require(igraph)

g <- graph_from_adjacency_matrix(co, 
                                 weighted=TRUE,     # weighted graph
                                 mode="undirected", # no direction
                                 diag=FALSE         # no looping edges
                                )

# Now we plot
plot.igraph(g, 
            layout=layout_with_gem,       # how to lay out the graph
            vertex.label.cex=0.6,         # label size
            vertex.label.color="black",   # font color
            vertex.color=qd.colors[4],    # vertex color
            vertex.frame.color=NA,        # no vertex border
            edge.width=E(g)$weight        # scale edges according to their weight
            )  
```

## Independent Vertex Sets

Independent vertex sets are sets of characters that do not have edges between any two characters. I.e., there is not a single scene in the play in which the characters shown in the vertex sets below are co-present. This is based on the function [`largest_ivs()`](https://igraph.org/r/doc/ivs.html) from the igraph package.

```{r}
largest_ivs(g)
```


## Gossip Network

The "gossip network" represents character that talk about other characters. The colors don't follow a particular pattern and are just added for readability. Edge thickness represent how often (number of scenes) this happens, and is scaled logarithmically.
Currently, this includes scenes in which both characters are present. This will be changed later.

```{r}
m <- as.matrix(
  as.data.frame.matrix(
    xtabs( ~ Speaker.figure_id + Mentioned.figure_id,
           mtext[,.N,.(Speaker.figure_id,Mentioned.figure_id,begin.Scene)]
           [!is.na(Mentioned.figure_id)])))

g <- graph_from_adjacency_matrix(m, 
                                 weighted=TRUE,     # weighted graph
                                 mode="directed", # no direction
                                 diag=FALSE         # no looping edges
                                )
coords <- layout_(g, on_grid())

plotGossipNetwork <- TRUE
if (length(igraph::E(g)) == 0) {
  plotGossipNetwork <- FALSE
}
```

```{r, eval=plotGossipNetwork}

plot.igraph(g,
            layout=coords,                # how to lay out the graph
            vertex.label.cex=0.6,         # label size
            vertex.label.color="black",   # font color
            vertex.color=qd.colors[4],    # vertex color
            vertex.frame.color=NA,        # no vertex border
            edge.curved = TRUE,           # curved edges
            edge.width=log(E(g)$weight)+1,  # scale edges according to their weight
            edge.color=rep(qd.colors,100)
            )  

```


# Word Fields  {.tabset}

This analysis shows word fields only for characters that speak more than `r params$minTokens` tokens. This value can be adjusted by supplying the argument `minTokens` to the `report()` function. Z-Sore analysis is only shown if more than one character remains after this filtering, other plots only if at least one character remains.

## Bar Chart

```{r, echo=FALSE}
showZScores <- TRUE

baseUrl <- "https://raw.githubusercontent.com/quadrama/metadata/ec8ae3ddd32fa71be4ea0e7f6c634002a002823d/fields/"

fieldNames <- c("Liebe", "Krieg", "Familie", "Ratio", "Religion")
fields <- loadFields(fieldNames,baseurl = baseUrl)

text2 <- limitFigures(mtext, by="tokens", threshold=params$minTokens)
dstat <- dictionaryStatistics(
  text2,  # the text
  fields=fields,
  names = TRUE,                 # use figure names (instead of ids)
  normalizeByFigure = TRUE,   # normalization by figure
  normalizeByField = TRUE,    # normalization by field
  column = "Token.lemma",        # lemma-based stats
  asList = TRUE
)
rownames(dstat$mat) <- toFigureName(rownames(dstat$mat))
par(mar=c(10,2,0,10),xpd=TRUE)
b <- barplot(t(dstat$mat), 
        col=params$col[1:length(fields)],
        las=3,
        border=FALSE
        )
legend(x=max(b)+1,y=max(dstat$mat),
       legend=colnames(dstat$mat),
       fill=params$col,
       bty="n",
       border=FALSE)

```

## Bar chart (z-scores)


```{r, results="asis"}
if (length(dstat$figure) < 2) {
  message("*Less than two characters remain, skipping Z-Score analysis.*")
  showZScores <- FALSE
}
```

[Z-scores in wikipedia](https://en.wikipedia.org/wiki/Standard_score)

```{r echo=FALSE, eval=showZScores}
zmat <- scale(dstat$mat)


par(mar=c(2,12,0,6),xpd=TRUE)
b <- barplot(t(zmat), 
        col=params$col[1:length(fieldNames)],
        las=1,horiz=TRUE,
        border=FALSE,beside=TRUE
        )
legend(x=max(zmat),y=max(b)+1,legend=colnames(zmat),fill=params$col,bty="n",border=FALSE)

```


## Bar chart (scaled)

```{r, echo=FALSE}
mat <- t(scale(t(dstat$mat),center=FALSE,scale=rowSums(dstat$mat)))

par(mar=c(10,2,1,10),xpd=TRUE)
b <- barplot(t(mat),
        col=params$col[1:length(fields)],
        las=3,
        border=FALSE
        )
legend(x=max(b)+1,y=max(dstat$mat),legend=colnames(dstat$mat),fill=params$col,bty="n",border=FALSE)

```


## Spider Web

```{r, echo=FALSE}
plotSpiderWebs(dstat, col=params$col)
```

## Word fields

```{r, echo=FALSE}
fields
```

# Personnel Exchange   {.tabset}

```{r, echo=FALSE}
showPersonnelExchange = length(unique(mtext$begin.Act)) > 1
```

```{r,echo=FALSE,eval=!showPersonnelExchange, results='asis'}
print("*Not applicable for this file.*")
```


## Static

The following chart shows for each scene boundary the number of characters exchanged over the boundary. Different metrics have been proposed, the plot shows all of them in comparison.


```{r, echo=FALSE, eval=showPersonnelExchange}
df <- list()
df$ham <- hamming(mtext, variant="NormalizedHamming")
df$trilcke <- hamming(mtext, variant="Trilcke")
df$sd <- scenicDifference(mtext)

if (length(df$ham)+1 != length(unique(mtext$begin.Scene))) {
  message("Something is seriously wrong here. Please report this issue.")
  showPersonnelExchange <- FALSE
}
```

```{r, echo=FALSE, eval=showPersonnelExchange}

df <- list()
df$ham <- hamming(mtext, variant="NormalizedHamming")
df$trilcke <- hamming(mtext, variant="Trilcke")
df$sd <- scenicDifference(mtext)



if (is.null(params$col)) {
  colors <- gray.colors(3)
} else {
  colors <- params$col[c(1,4,7)]
}

names <- c("Hamming (Normalized)", 
                             "Trilcke et al. (2017)", 
                             "Scenic Difference")
sceneLabels <- paste(as.roman(mtext[,.N,.(Number.Act, Number.Scene)]$Number.Act), mtext[,.N,.(Number.Act, Number.Scene)]$Number.Scene, sep=".")
  par(mar=c(3,3,3,3))
  layout(matrix(c(1,2,2,2,2),5,1,byrow=TRUE))
  plot(0,type="n",axes=FALSE,xlab="",ylab="")
  legend(x="center",legend=c("Hamming (Normalized)", 
                             "Trilcke et al. (2017)", 
                             "Scenic Difference"),bty="n",horiz=TRUE,fill = colors)
  
  
  actBoundaries <- mtext[,length(unique(begin.Scene)),.(begin.Act)]$V1
  s <- 0
  for (i in 1:length(actBoundaries)) {
    actBoundaries[i] <- actBoundaries[i] + s
    s <- actBoundaries[i]
  }
  
  b <- barplot(Reduce(rbind, df), beside=TRUE,col = colors, border=FALSE)
  axis(1, at=seq(0.5,max(b[3,])+1,4),labels=sceneLabels, cex=0.1)
  


```

## Dynamic

```{r, echo=FALSE, eval=showPersonnelExchange}
highchart() %>%
  hc_yAxis(max=1) %>%
  hc_xAxis(categories=sceneLabels,
           tickmarkPlacement="between",
           labels.x=-15) %>%
  hc_add_series(df$ham, pointPadding=0,
                type="column",
                name=names[[1]],
                color=colors[[1]]) %>%
  hc_add_series(df$trilcke,
                pointPadding=0,
                type="column",
                name=names[[2]],
                color=colors[[2]]) %>%
  hc_add_series(df$sd, pointPadding=0,
                type="column",
                name=names[[3]],
                color=colors[[3]])
```

# About

This report has geen generated using the R package `DramaAnalysis` in version `r pver`. The package is being developed [here](http://github.com/quadrama/DramaAnalysis), in the context of the project [QuaDramA](https://quadrama.github.io). Please report errors [here](https://github.com/quadrama/DramaAnalysis/issues), or contact [me](mailto:nils.reiter@ims.uni-stuttgart.de).

## Reproduction
For the generation of this report, the following list of parameters was used:

```{r}
params
```

The following table describes these options

| Name  | Meaning |
| ----- | ------- |
| `id`  | The id of the play we want to report on |
| `topN` | If an analysis only takes the most talkative characters into account, it will use the number of characters given here. Default: `10` |
| `minTokens` | If an analysis requires characters to speak at least a certain number of tokens, this value can be adjusted here. Default: `500` | 
| `col` | A list of color codes in a form R can understand. Default: ```r qd.colors``` | 

