---
title: "QuaDramA Report"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
  pdf_document:
    toc: true
    toc_depth: 2
params:
  id: "tg:rksp.0"
  col: "grey"
---

```{r,echo=FALSE}
pver <- packageVersion("DramaAnalysis")
```

This report has geen generated using the R package `DramaAnalysis` in version `r pver`. The package is being developed [here](http://github.com/quadrama/DramaAnalysis), in the context of the project [QuaDramA](https://quadrama.github.io).

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setup()

text <- loadText(params$id, includeTokens = TRUE)
mtext <- loadSegmentedText(params$id)

```

# Figure Lists {.tabset}

## Dramatis Personae 

All figurs as listed in the dramatis personae:

```{r, echo=FALSE}
figuresInDramatisPersonae <- loadAnnotations(params$id, type=atypes$DramatisPersonae, coveredType = atypes$Figure)
as.character(figuresInDramatisPersonae$Figure.surface)
```

## By Order of appearance

All figures, in the order they have their first appearance

```{r, echo=FALSE}
begins <- aggregate(text$begin, by=list(text$Speaker.figure_surface), min)
as.character(begins[order(begins$x),]$Group.1)
```

# Overall Speech Distribution

Relative amount of words spoken by each figure. 

```{r, echo=FALSE}
fstat <- figureStatistics(text, names=TRUE, normalize=FALSE)
mat <- figurematrix(fstat)
b <- barplot(mat$values,col=params$col)

top <- 15

text(x=b, y=t(head(mat$cs,top)+(head(mat$values,top)/2)),
     labels=t(substr(head(mat$labels,top),0,20)),cex=0.5)
```

# Presence and Utterances

```{r,echo=FALSE}
ustat <- utteranceStatistics(mtext, normalizeByDramaLength = FALSE, numberOfFigures = FALSE)

```

## All Utterances
```{r,echo=FALSE}
par(mar=c(1,9,1,0),xpd=FALSE)
plotUtterancePositions(ustat,mtext,xlab="")
```

## Utterance Length Variation

```{r,echo=FALSE}
par(mar=c(2,9,0,0))
boxplot(utteranceLength ~ figure, data=ustat, horizontal=TRUE,las=1,frame=FALSE)
```

# Configuration and Copresence {.tabset}

## By Act 
```{r, echo=FALSE}
c <- configuration(mtext)
c$matrix <- scale(c$matrix, center=FALSE, scale=colSums(c$matrix))
barplot(c$matrix, 
        legend.text=c$figure, # set legend text
        args.legend = list(cex=0.5, # legend font size
                           x=7.5, # legend x position
                           y=max(colSums(c$matrix)) # legend y pos
                        ),
        col=params$col)
```

## By Scene

```{r, echo=FALSE}
c <- configuration(mtext, by="Scene")
c$matrix <- scale(c$matrix, center=FALSE, scale=colSums(c$matrix))
barplot(c$matrix, 
        legend.text=c$figure, # set legend text
        args.legend = list(cex=0.5, # legend font size
                           x=7.5, # legend x position
                           y=max(colSums(c$matrix)) # legend y pos
                        ),
        col=params$col)
```


# Copresence Network
```{r, echo=FALSE, message=FALSE}
c <- configuration(mtext, onlyPresence = TRUE, by="Scene")
co <- c$matrix %*% t(c$matrix)

# add figure names
rownames(co) <- c$figure
colnames(co) <- c$figure
require(igraph)

g <- graph_from_adjacency_matrix(co, 
                                 weighted=TRUE,     # weighted graph
                                 mode="undirected", # no direction
                                 diag=FALSE         # no looping edges
                                )

# Now we plot
plot.igraph(g, 
            layout=layout_with_gem,       # how to lay out the graph
            vertex.label.cex=0.6,         # label size
            vertex.label.color="black",   # font color
            vertex.color=qd.colors[4],    # vertex color
            vertex.frame.color=NA,        # no vertex border
            edge.width=E(g)$weight        # scale edges according to their weight
            )  
```


# Word Fields  {.tabset}

## Bar Chart

```{r, echo=FALSE}
baseUrl <- "https://raw.githubusercontent.com/quadrama/metadata/ec8ae3ddd32fa71be4ea0e7f6c634002a002823d/fields/"

text2 <- limitFigures(text, by="tokens")
fields <- c("Liebe", "Krieg", "Familie", "Ratio", "Religion")
dstat <- dictionaryStatistics(
  text2,  # the text
  fieldnames=fields,  # fields we want to measure
  baseurl = baseUrl,            # specify url
  names = TRUE,                 # use figure names (instead of ids)
  normalizeByFigure = TRUE,   # normalization by figure
  normalizeByField = TRUE,    # normalization by field
  column = "Token.lemma"        # lemma-based stats
)
mat <- as.matrix(dstat[,3:7])
rownames(mat) <- vapply(strsplit(as.character(dstat$figure), ",", fixed=TRUE), `[`,1, FUN.VALUE=character(1))

par(mar=c(10,2,0,10),xpd=TRUE)
b <- barplot(t(mat), # we select Romeo's line
        col=params$col[1:length(fields)],
        las=3,
        border=FALSE
        )
legend(x=max(b)+1,y=max(mat),legend=colnames(mat),fill=params$col,bty="n",border=FALSE)

```

## Bar chart (scaled)

```{r, echo=FALSE}
mat <- t(scale(t(mat),center=FALSE,scale=rowSums(mat)))

par(mar=c(10,2,1,10),xpd=TRUE)
b <- barplot(t(mat),
        col=params$col[1:length(fields)],
        las=3,
        border=FALSE
        )
legend(x=max(b)+1,y=max(mat),legend=colnames(mat),fill=params$col,bty="n",border=FALSE)

```


## Spider Web

```{r, echo=FALSE}
plotWordFields(dstat, col=params$col)
```